---
title: "Statistical Analysis in R – CVD & Customer Satisfaction Report"
author: 'Joao Marques Hassun'
date: "2024-12-06"
output: html_document
---

# Load the libraries.

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(gridExtra)
library(kableExtra)
library(emmeans) # for emmeans() and pairs()
library(Hmisc)
library(car) # for vif()
options(width=100)

```

---

## First Request

## Data Dictionary

The variables are described in the table below:

Variable | Description
------------- | -------------
area_name| The name of the local authority region in which the data was captured.
area_code| The code of the local authority region in which the data was captured.
Population| The total population of the local authority region in which the data was captured.
Poverty| The proportion who meet the definition of living in poverty.
CVD| the percentage of people living in the area who have recently experienced Cardiovascular Disease (CVD).
overweight| The proportion who are overweight.
smokers| The proportion who smoke.
wellbeing| The average wellbeing score of people living in the area.

---

# Read the csv file and assign it to a variable.

```{r}

S1 <- read_csv('Cardio_Vascular_Disease.csv')

```

# Check data integrity.

```{r}

summary(S1)

# running a summary on the dataset reveals 76 NAs in Population, 76 NAs in Poverty, 76 NAs in CVD, 72 NAs in Overweight, 7 NAs in Smokers and 15 NAs in Wellbeing.

```

# Remove NAs.

```{r}

# To ensure the integrity of the dataset and avoid any chance of bias or distorted results, 82 rows with missing values were removed.

S1 <- S1 %>% drop_na()

# Run the summary again to ensure that changes have taken place.

summary(S1)

```

# Standardize column names.

```{r}

colnames(S1) <- c("Area Name", "Area Code", "Population", "Poverty", "CVD", "Overweight", "Smokers", "Wellbeing")

# Check that changes have taken place.

colnames(S1)

```

# Visualise data to check data interactions and distribution.

```{r}

ggplot(S1, aes(x = Poverty, y = CVD)) + geom_point(color = "blue", alpha = 0.5) + geom_smooth(method = "lm", color = "red") + labs(title = "Effect of Poverty upon CVD", x = "Poverty", y = "CVD") + theme(plot.title = element_text(hjust = 0.5))

# The Poverty-CVD scatter plot suggests a negative effect of Poverty upon CVD.

ggplot(S1, aes(x = Overweight, y = CVD)) + geom_point(color = "blue", alpha = 0.5) + geom_smooth(method = "lm", color = "red") + labs(title = "Effect of Overweight upon CVD", x = "Overweight", y = "CVD") + theme(plot.title = element_text(hjust = 0.5))

# The Overweight-CVD scatter plot suggests a positive effect of Poverty upon CVD.

ggplot(S1, aes(x = Smokers, y = CVD)) + geom_point(color = "blue", alpha = 0.5) + geom_smooth(method = "lm", color = "red") + labs(title = "Effect of Smokers upon CVD", x = "Smokers", y = "CVD") + theme(plot.title = element_text(hjust = 0.5))

# The Smokers-CVD scatter plot suggests a positive effect of Poverty upon CVD.

ggplot(S1, aes(x = Wellbeing, y = CVD)) + geom_point(color = "blue", alpha = 0.5) + geom_smooth(method = "lm", color = "red") + labs(title = "Effect of Wellbeing upon CVD", x = "Wellbeing", y = "CVD") + theme(plot.title = element_text(hjust = 0.5))

# The Wellbeing-CVD scatter plot suggests a positive effect of Poverty upon CVD.

ggplot(S1, aes(x = CVD)) + geom_histogram(color = "black", fill = "blue", alpha = 0.5, bins = 100) + labs(title = "CVD Histogram", x = "CVD Prevalence", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5))

# CVD data appears to be normally distributed.

ggplot(S1, aes(x = Poverty)) + geom_histogram(color = "black", fill = "blue", alpha = 0.5, bins = 100) + labs(title = "Poverty Histogram", x = "Poverty Prevalence", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5))

# Poverty data appears to have slight positive skew, but close to normal.

ggplot(S1, aes(x = Overweight)) + geom_histogram(color = "black", fill = "blue", alpha = 0.5, bins = 100) + labs(title = "Overweight Histogram", x = "Overweight Prevalence", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5))

# Overweight data appears to be normally distributed.

ggplot(S1, aes(x = Smokers)) + geom_histogram(color = "black", fill = "blue", alpha = 0.5, bins = 100) + labs(title = "Smokers Histogram", x = "Smokers Prevalence", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5))

# Smokers data appears to be normally distributed.

ggplot(S1, aes(x = Wellbeing)) + geom_histogram(color = "black", fill = "blue", alpha = 0.5, bins = 100) + labs(title = "Wellbeing Histogram", x = "Wellbeing Prevalence", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5))

# Wellbeing data appears to be normally distributed.

# All histograms (CVD, Poverty, Overweight, Smokers, and Wellbeing) suggest the possibility of outliers.

```

# Create a model to analyse which factors impact the prevalence of CVD in areas.

```{r}

M1 <- lm(CVD ~ Poverty + Overweight + Smokers + Wellbeing, data = S1)

# Create a summary of the model.

summary(M1)

# The intercept is not statistically significant (t(298) = -0.432, p>0.05) which suggests that when all predictors are zero the baseline level of CVD is not significantly different from zero.

# There is a statistically significant negative effect of Poverty upon CVD (t(298) = -5.234, p<0.05), indicating that higher levels of Poverty are associated with lower levels of CVD.

# There is a statistically significant positive effect of Overweight upon CVD (t(298) = 5.174, p<0.05), indicating that higher levels of Overweight are associated with higher levels of CVD.

# There is a statistically significant positive effect of Smokers upon CVD (t(298) = 3.574, p<0.05), indicating that higher levels of Smokers are associated with higher levels of CVD.

# There is a statistically significant positive effect of Wellbeing upon CVD (t(298) = 3.667, p<0.05), indicating that higher levels of Wellbeing are associated with higher levels of CVD.

```

# Run confint to check the confidence intervals of the model.

```{r}

confint(M1)

# With every 1 point increase in Poverty, the model predicts an average decrease of 0.18400 points on CVD with 95% CI, (CI = [-0.25, -0.11]).

# With every 1 point increase in Overweight, the model predicts an average increase of 0.10985 points on CVD with 95% CI, (CI = [0.06, 0.15]).

# With every 1 point increase in Smokers, the model predicts an average increase of 0.12030 points on CVD with 95% CI, (CI = [0.05, 0.18]).

# With every 1 point increase in Wellbeing, the model predicts an average increase of 1.80025 points on CVD with 95% CI, (CI = [0.83, 2.76]).

```

# Run RCorr to explore the relationships between variables.

```{r}

rcorr(as.matrix(select_if(S1, is.numeric)))

# rcorr suggests that there is a statistically significant positive relationship between Population and Poverty (r = 0.33, p<0.05), indicating that higher levels of Population are associated with higher levels of Poverty.

# rcorr suggests that there is a statistically significant negative relationship between Population and CVD (r = -0.20, p<0.05), indicating that higher levels of Population are associated with lower levels of Poverty.

# rcorr suggests that the relationship between Population and Overweight is not statistically significant (r = -0.02, p>0.05).

# rcorr suggests that the relationship between Population and Smokers is not statistically significant (r = 0.06, p>0.05).

# rcorr suggests that there is a statistically significant negative relationship between Population and Wellbeing (r = -0.21, p<0.05), indicating that higher levels of Population are associated with lower levels of Wellbeing

# rcorr suggests that there is a statistically significant negative relationship between Poverty and CVD (r = -0.25, p<0.05), indicating that higher levels of Poverty are associated with lower levels of CVD.

# rcorr suggests that there is a statistically significant negative relationship between Poverty and Overweight (r = -0.02, p<0.05), indicating that higher levels of Poverty are associated with lower levels of Overweight.

# rcorr suggests that there is a statistically significant positive relationship between Poverty and Smokers (r = 0.36, p<0.05), indicating that higher levels of Poverty are associated with higher levels of Smokers.

# rcorr suggests that there is a statistically significant negative relationship between Poverty and Wellbeing (r = -0.35, p<0.05), indicating that higher levels of Poverty are associated with lower levels of Wellbeing.

# rcorr suggests that there is a statistically significant positive relationship between CVD and Overweight (r = 0.32, p<0.05), indicating that higher levels of CVD are associated with higher levels of Overweight.

# rcorr suggests that there is a statistically significant positive relationship between CVD and Smokers (r = 0.18, p<0.05), indicating that higher levels of CVD are associated with higher levels of Smokers.

# rcorr suggests that there is a statistically significant positive relationship between CVD and Wellbeing (r = 0.25, p<0.05), indicating that higher levels of CVD are associated with higher levels of Wellbeing.

# rcorr suggests that there is a statistically significant positive relationship between Smokers and Overweight (r = 0.40, p<0.05), indicating that higher levels of Smokers are associated with higher levels of Overweight.

# rcorr suggests that there is a statistically significant negative relationship between Smokers and Wellbeing (r = -0.20, p<0.05), indicating that higher levels of Smokers are associated with lower levels of Wellbeing.

# rcorr suggests that the relationship between Overweight and Wellbeing is not statistically significant (r = -0.04, p>0.05).

```

# Run Variance Inflation Factor (VIF) to check for multicollinearity and analyse the relationship between predictors.

```{r}

vif(M1)

# Considering that all VIF values are low (<5), we can state that multicollinearity is not significant in this case.

```

# Create a figure to visualise the effect of poverty upon CVD.

```{r}

ggplot(S1, aes(x = Poverty, y = CVD)) + geom_point(color = "blue", alpha = 0.5) + geom_smooth(method = "lm", color = "red") + labs(title = "Effect of Poverty upon CVD", x = "Poverty", y = "CVD") + theme(plot.title = element_text(hjust = 0.5))

```

---

## Second Request

## Data Dictionary

The variables are described in the table below:

Variable | Description
------------- | -------------
SES_category| The company's categorisation of store type by local socio-economic-status (low, medium, high).
customer.satisfaction| The average customer satisfaction score.
staff.satisfaction| The average staff job satisfaction score.
delivery.time| The average delivery time of large and custom items.
new_range| Whether the store was carrying a new range of products or not (true, false).

---

# Read the csv file and assign it to a variable.

```{r}

S2 <- read_csv('cust_satisfaction.csv')

```

# Run a summary to check data integrity.

```{r}

# The summary reveals no NAs present.

summary(S2)

```

# Standardize column names.

```{r}

colnames(S2) <- c("SES Category", "Customer Satisfaction", "Staff Satisfaction", "Delivery Time", "New Range")

# Check that changes have taken place

colnames(S2)

```

# Visualize data distribution and interaction between factors.

```{r}

ggplot(S2, aes(x = `Customer Satisfaction`)) + geom_histogram(fill = "blue", color = "black", bins = 100) + labs(title = "Customer Satisfaction Histogram", x = "Customer Satisfaction Prevalence", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5))

# The Customer Satisfaction Histogram appears to have a slight negative skew, but close to normal.

ggplot(S2, aes(x = `Staff Satisfaction`)) + geom_histogram(fill = "blue", color = "black", bins = 100) + labs(title = "Staff Satisfaction Histogram", x = "Staff Satisfaction Prevalence", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5))

# The Staff Satisfaction Histogram appears to have a slight positive skew, but close to normal.

ggplot(S2, aes(x = `Delivery Time`)) + geom_histogram(fill = "blue", color = "black", bins = 100) + labs(title = "Delivery Time Histogram", x = "Delivery Time Prevalence", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5))

# The Delivery Time Histogram appears to have a slight positive skew, but close to normal.

ggplot(S2, aes(x = `Staff Satisfaction`, y = `Customer Satisfaction`)) + geom_point(color = "blue", alpha = 0.5) + geom_smooth(method = "lm", color = "red") + labs(title = "Effect of Staff Satisfaction upon Customer Satisfaction", x = "Staff Satisfaction", y = "Customer Satisfaction") + theme(plot.title = element_text(hjust = 0.5))

# The Effect of Staff Satisfaction upon Customer Satisfaction scatter plot suggests a positive effect of Staff Satisfaction upon Customer Satisfaction.

ggplot(S2, aes(x = `Delivery Time`, y = `Customer Satisfaction`)) + geom_point(color = "blue", alpha = 0.5) + geom_smooth(method = "lm", color = "red") + labs(title = "Effect of Delivery Time upon Customer Satisfaction", x = "Delivery Time", y = "Customer Satisfaction") + theme(plot.title = element_text(hjust = 0.5))

# The Effect of Delivery Time upon Customer Satisfaction scatter plot suggests a negative effect of Staff Satisfaction upon Customer Satisfaction.

ggplot(S2, aes(x = `SES Category`, y = `Customer Satisfaction`, fill = `New Range`)) + geom_boxplot(color = "blue", alpha = 0.5) + labs(title = "Effect of SES Category and New Range upon Customer Satisfaction", x = "SES Category", y = "Customer Satisfaction", fill = "New Range") + theme(plot.title = element_text(hjust = 0.5))

# There appears to be an outlier in the data of High SES Stores which offer a New Range of products.

```

# Create a model to analyse which factors impact Customer Satisfaction.

```{r}

M2 <- lm(`Customer Satisfaction` ~ `Staff Satisfaction` + `Delivery Time` + `SES Category` + `New Range`, data = S2)

# Create a summary of the model.

summary(M2)

# The intercept is statistically significant (t(294) = 8.550, p<0.05) which suggests that when all predictors are zero the baseline level of customer satisfaction is significantly different from zero.

# There is a statistically significant positive effect of Staff Satisfaction upon Customer Satisfaction (t(294) = 4.364, p<0.05), indicating that higher levels of staff satisfaction are associated with higher levels of customer satisfaction.

# There is a statistically significant negative effect of Delivery Time upon Customer Satisfaction (t(294) = -3.480, p<0.05), indicating that higher delivery times are associated with lower levels of customer satisfaction.

# There is no statistically significant effect of Low SES Category stores upon Customer Satisfaction (t(294) = -1.846, p>0.05).

# There is a statistically significant positive effect of Medium SES Category stores upon Customer Satisfaction (t(294) = 8.183, p<0.05), indicating that Medium SES Category stores are associated with higher levels of customer satisfaction.

# There is no statistically significant effect of stores carrying a new range of products upon Customer Satisfaction (t(294) = 0.836, p>0.05).

```

# Create a model to analyse how all factors affect customer satisfaction, particularly how the effect of Delivery Time on Customer Satisfaction interacts with SES categories.

```{r}

M3 <- lm(`Customer Satisfaction` ~ `Staff Satisfaction` + `Delivery Time` * `SES Category` + `New Range`, data = S2)

# Create a summary of the model.

summary(M3)

# The intercept is statistically significant (t(292) = 8.380, p<0.05) which suggests that when all predictors are zero the baseline level of customer satisfaction is significantly different from zero.

# There is a statistically significant positive effect of Staff Satisfaction upon Customer Satisfaction (t(292) = 4.504, p<0.05), indicating that higher levels of staff satisfaction are associated with higher levels of customer satisfaction.

# There is a statistically significant negative effect of Delivery Time upon Customer Satisfaction (t(292) = -3.960, p<0.05), indicating that higher delivery times are associated with lower levels of customer satisfaction.

# There is a statistically significant negative effect of Low SES Category stores upon Customer Satisfaction (t(292) = -2.964, p<0.05), indicating that Low SES category stores are associated with lower customer satisfaction.

# There is no statistically significant effect of Medium SES Category stores upon Customer Satisfaction (t(292) = -0.002, p>0.05).

# There is no statistically significant effect of stores carrying a new range of products upon Customer Satisfaction (t(292) = 0.35103, p>0.05).

# There is a statistically significant interaction between Delivery Time and Low SES category stores (t(292) = 2.670, p<0.05), indicating that the negative effect of longer delivery times on Customer Satisfaction decreases for Low SES stores. 

# In contrast, the interaction between Delivery Time and Medium SES category stores is not statistically significant (t(292) = 1.625, p>0.05).

```

# Run ANOVA to compare the two models (M2 and M3).

```{r}

anova(M2, M3)

# ANOVA reveals that, as the p-value (p=2.2e-16) is statistically significant (p<0.05), we can reject the null hypothesis and can also affirm that the more complex model (M3) is better than the simpler model (M2).

```


# Run Emmeans to analyse the interaction between Delivery Time and SES Category in M3.

```{r}

emmeans_M3 <- emmeans(M3, ~ `Delivery Time`*`SES Category`)
emmeans_M3

# For a delivery time of 59.6, the emmeans of Delivery Time varies across high, medium and low SES stores. The results indicate that medium SES stores achieve a higher level of customer satisfaction (emmean = 7.82), in comparision to High (emmean = 6.61) and Low (emmean = 6.32) SES category stores. 

```

# Create a model focusing solely on how Customer Satisfaction is affected by Delivery Time and SES Categories.

```{r}

M4 <- lm(`Customer Satisfaction` ~ `Delivery Time` + `SES Category`, data = S2)
summary(M4)

# The intercept is statistically significant (t(296) = 23.705, p<0.05) which suggests that when all predictors are zero the baseline level of customer satisfaction is significantly different from zero.

# There is a statistically significant negative effect of Delivery Time upon Customer Satisfaction (t(296) = -3.329, p<0.05), indicating that higher delivery times are associated with lower levels of customer satisfaction.

# There is a statistically significant negative effect of Low SES Category stores upon Customer Satisfaction (t(296) = -2.231, p<0.05), indicating that Low SES Category stores are associated with lower levels of customer satisfaction.

# There is a statistically significant positive effect of Medium SES Category stores upon Customer Satisfaction (t(296) = 10.199, p<0.05), indicating that Medium SES Category stores are associated with higher levels of customer satisfaction.

```

# Create a model focusing solely on how Customer Satisfaction is affected by the interaction between Delivery Time and SES Categories.

```{r}

M5 <- lm(`Customer Satisfaction` ~ `Delivery Time` * `SES Category`, data = S2)
summary(M5)

# The intercept is statistically significant (t(294) = 15.072, p<0.05) which suggests that when all predictors are zero the baseline level of customer satisfaction is significantly different from zero.

# There is a statistically significant negative effect of Delivery Time upon Customer Satisfaction (t(294) = -3.669, p<0.05), indicating that higher delivery times are associated with lower levels of customer satisfaction.

# There is a statistically significant negative effect of Low SES Category stores upon Customer Satisfaction (t(294) = -2.740, p<0.05), indicating that Low SES Category stores are associated with lower levels of customer satisfaction.

# There is no statistically significant effect of Medium SES Category stores upon Customer Satisfaction (t(294) = 0.388, p>0.05).

# There is a statistically significant interaction between Delivery Time and Low SES category stores (t(294) = 2.374, p<0.05), indicating that the negative effect of longer delivery times on Customer Satisfaction decreases for Low SES stores.

# There is no statistically significant interaction between Delivery Time and Medium SES category stores (t(294) = 1.505, p>0.05). 

```

# Run ANOVA to compare the two models (M4 and M5).

```{r}

anova(M4, M5)

# ANOVA reveals that, as the p-value (p=0.05977) is not statistically significant (p>0.05), we cannot reject the null hypothesis and can also affirm that the simpler model (M4) is better than the more complex model (M5).

```

# Run Emmeans to analyse the interaction between Delivery Time and SES Category in M5.

```{r}

emmeans_M5 <- emmeans(M5, ~ `Delivery Time`*`SES Category`)
emmeans_M5

# For a delivery time of 59.6, the emmeans of Delivery Time varies across high, medium and low SES stores. The results indicate that medium SES stores achieve a higher level of customer satisfaction (emmean = 7.80), in comparision to High (emmean = 6.36) and Low (emmean = 6.00) SES category stores. 

```

# Visualise the interaction between Customer Satisfaction and Delivery Time, and whether the effect is the same across high, medium and low ‘’Socio-Economic-Status’’ (SES) stores.

```{r}

ggplot(S2, aes(x = `Delivery Time`, y = `Customer Satisfaction`, color = `SES Category`)) + geom_point() + geom_smooth(method = "lm") + labs(title = "Interaction of Delivery Time and Customer Satisfaction across SES Categories", x = "Delivery Time", y = "Customer Satisfaction", color = "Store Type") + theme(plot.title = element_text(hjust = 0.5))

# The first scatter plot displays all SES Categories in one single figure, which allows for an overall view of trends and interactions. Color is used to distinguish between categories.

ggplot(S2, aes(x = `Delivery Time`, y = `Customer Satisfaction`, color = `SES Category`)) + geom_point() + geom_smooth(method = "lm") + facet_grid(`SES Category`~.) + labs(title = "Interaction of Delivery Time and Customer Satisfaction by SES Category", x = "Delivery Time", y = "Customer Satisfaction", color = "Store Type") + theme(plot.title = element_text(hjust = 0.5))

# The second scatter plot uses facet_grid() to separate the data into separate panels for each SES Category, which helps compare each category.

```

